ICT and Future Productivity:\\Evidence and Theory of a GPT\thanks{Correspondence: Department of Economics, Boston College, 140 Commonwealth Avenue, Chestnut Hill, MA 02467. Email: brianti@bc.edu (Marco Brianti) and gati@bc.edu (Laura G\'ati).}

### 2018-07-07 Latex/Draft_Summer2018/Summer2018.tex
starting the model

@@

>>>>

Figures \ref{fig:2stepIRF_first} and \ref{fig:2stepIRF_second} in Appendix \ref{section:IRF_2step} shows estimated impulse responses to both news and ICT shocks. In particular, Figure \ref{fig:2stepIRF_first} shows dynamic responses of TFP, stock prices, and ICT investment to both shocks and Figure \ref{fig:2stepIRF_second} shows dynamic responses of GDP, consumption and relative prices to both shocks as well. The shaded gray areas are the $90$\% and $95$\% confidence bands from a bias-corrected bootstrap procedure of \cite{kilian1998small} using 2000 simulations.

Dynamic responses to an ICT shock are almost identical to the ones presented in the main identification in Section \ref{section:mainSpec}. After few lags, TFP displays a positive and significant effect and reaches its peak of $1.2$\% after $6$ years. Interestingly, stock prices jump on impact and the effect remain significant over the medium run. Real ICT investment has a large and positive impact response and it slowly starts to decay remaining significant for a long period of time. Both real GDP and real consumption have a significant impact response and reaches its peak approximately after $6$ years. Finally, the response of relative prices is negative and persistent as we have previously discussed.

In addition, we show dynamic responses to a news shock to confirm the reliability of our identification strategy. Constraining fluctuations of relative prices to news does not qualitatively affect what has been already found in the literature. In line with \cite{barsky2011news}, it takes more that 2 years for TFP to display a positive, significant and strongly persistent effect. Stock prices have a strong forecasting power and show a $4$\% impact effect which is persistent and significant in the long run. Real ICT investment does not show a positive impact response but it slowly starts to increase getting significant after almost $2$ years. As found by \cite{barsky2011news} in their main specification, real GDP has a non-significant impact effect and it fast rises with all the variables of the system. Moreover, as expected by its forward looking nature, real consumption have a significant impact response and reaches its peak once TFP displays a significant effect. Finally, relative prices are initially constrained to have a zero response and then shows a small, negative and persistent deviation for at least five years which eventually dies out.

Finally, we recover a series of ICT shocks which is correlated at $94$\% with the one estimated in System \ref{eq:mainSpecification} using the simple identification strategy presented in Section \ref{section:empiricalstrategy_simple}. In particular, in Appendix \ref{section:ICTshockseries_2step}, Figure \ref{fig:shockSeries_2steps} shows ICT shock series for both the 5-dimensional system and 6-dimensional presented in this section. In line with the correlation coefficient, the two series almost overlap each others suggesting that our initial result is not confounding any signal regarding future TFP with our estimated ICT shock.

%Here there is room for further comments:
%Variance explained by a news shock is only 20% -> smaller than the one in the literature. Maybe we should explain that news literature is confounding ICT shock into news shocks. However this is not really the target of our paper. Is it necessary to do? We need to think about that.

@@

In the previous section, we have documented results regarding the relation of current ICT investment and future productivity. Clearly, to rationalize empirical dynamic responses of aggregate macroeconomic variables to ICT shocks, we need a two-sector approach that distinguishes ICT goods from other output. In this section, we present a simple two-sector model that rationalizes our structural VAR estimates. The model presented has several common features with \cite{greenwood1997long} that design a two-sector economy where one sector displays a faster technology than the other. One different is that we focus specifically on ICT capital sector, rather than equipment, as the sector with faster technological improvements. This assumption will be necessary if we want to fit the fact that relative prices between the ICT investment and overall economy display a downward sloping trend.\footnote{In addition, we also rely on the results of \cite{whelan2003two} and \cite{oulton2007investment} to correctly define productivity and output share in our setting.}

\subsubsection*{Household}

Consider an economy inhabitated by a representative agent whi maximizes the expect values of his lifetime utility as given by
$$
\E \Bigg[  \sum_{t=0}^{\infty} \beta^t U(c_t,l_t)  \Bigg]
$$
with 
$$
U(c_t,l_t) = \log(c_t) - \frac{1}{1+\eta} l_t^{1+\eta}
$$
where $c_t$ and $l_t$ represent consumption and labor at time $t$.

\subsection*{Final-Good Sector}

In each period, the production of final output $y^c_t$ requires the services of labor $l_{1,t}$, and two types of capital: hard capital, $k^c_{1,t}$ and ICT capital, $k^i_{1,t}$.\footnote{Notice that $x_{1,t}$ has to be interpret as the part of overall input $x_t$ which is allocated to the final-output sector, i.e. sector $1$.} In addition, production function is augmented with (i) sector-specific productivity, $\Gamma^c_t$ and (ii) a positive spillover effect related to the diffusion of ICT capital in the whole economy, $k^i_t$.\footnote{Notice that differently to $k^i_{1,t}$, $k^i_t$ represents the total amount of ICT capital in the whole economy at time $t$.} Production of final output is undertaken in accordance with
$$
y^c_t = A^c_t \ (k^c_{1,t})^a \ (k^i_{1,t})^b \ (l_{1,t})^{1-a-b}
$$
$\Gamma^c_t k^i_{t} $
